---
title: "p8105_hw3_cz2955"
output: github_document
name: Chenhao Zheng
Uni: cz2955
date: "2025-10-06"
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(tidyverse)
library(lubridate)
library(patchwork)
library(ggplot2)
library(janitor)
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

### Load data:

```{r}
library(p8105.datasets)
data("instacart")
```

This dataset comprises partial information from Instacart order data, containing users' historical orders and itemised details of goods within those orders. Each row in the data represents a specific item purchased by a user during a particular order. Key variables include `r paste(colnames(instacart), collapse = ", ")`. There are total `r nrow(instacart)`variables. The average reorder rate across all items is `r round(mean(instacart$reordered, na.rm = TRUE), 3)`, indicating that a significant fraction of purchases are repeat orders. \### (a):

```{r}
length(unique(instacart$aisle))

aisle_counts <- instacart%>%
  group_by(aisle)%>%
  summarise(num_orders = n())%>%
  arrange(desc(num_orders))

head(aisle_counts)
```

### (b):

```{r}
aisle_counts_large <- aisle_counts %>%
  filter(num_orders > 10000)
ggplot(aisle_counts_large, aes(x = reorder(aisle, -num_orders), y = num_orders)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Number of Items Ordered by Aisle",
       x = "Aisle",
       y = "Number of Items Ordered") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### (c):

```{r}
popular_items <- instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarise(num_orders = n()) %>% 
  arrange(aisle, desc(num_orders)) %>%
  slice_head(n = 3)
  
popular_items
```

### (d)

```{r}
mean_order_hour <- instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>% 
  summarise(mean_hour = mean(order_hour_of_day, na.rm = TRUE)) %>% 
  pivot_wider(names_from = order_dow, values_from = mean_hour) %>% 
  arrange(product_name)

mean_order_hour
```

## Problem 2

```{r}
# Table1
Zip_Codes<-read_csv("Zip_Codes.csv")|>
  select(County, ZipCode, Neighborhood) |>
  mutate(
    ZipCode = as.integer(ZipCode),
    County = as.character(County),
    Neighborhood = as.character(Neighborhood)
    )
# Tablke2
Zip_NYC<-read_csv("Zip_NYC.csv")|>
  rename(ZipCode = RegionName,
         County = CountyName
         ) |>
  mutate(
    ZipCode = as.integer(ZipCode),
    County = str_replace(County, " County", "")
  )|>
  select(ZipCode, County, starts_with("201"), starts_with("202"))|>
  pivot_longer(
    cols = -c(ZipCode, County),
    names_to = "Date",
    values_to = "RentPrice"
  )
# Merge Table1, 2
z_together <- Zip_NYC |>
  left_join(Zip_Codes, by = c("ZipCode", "County"))

zip_counts <- z_together %>%
  drop_na(RentPrice) %>%
  group_by(ZipCode) %>%
  summarize(obs_count = n(), distinct_county = n_distinct(County)) %>%
  ungroup()

# Oberservation=116 times
n_116 <- zip_counts %>% 
  filter(obs_count == 116) %>% 
  nrow()

# Oberservation<10 times
n_lt10 <- zip_counts %>% 
  filter(obs_count < 10) %>% 
  nrow()
```

There are `r print(n_116)`ZIP codes observed 116 times and `r print(n_lt10)` ZIP codes are fewer than 10 times. Certain postcodes yield monthly data due to high rental activity and ample property listings within these areas; conversely, some postcodes show infrequent observations owing to small market size, infrequent listings, or incomplete data collection. Rental data is observed monthly for certain postcodes, typically because these areas feature active rental markets with ample supply and consistent data collection coverage.

### Table showing the average rental price

```{r}
# Table for average price
avg_rent <- z_together %>%
  mutate(Year = floor_date(as.Date(Date), unit = "year")) %>%  
  group_by(County, Year) %>%
  summarize(
    AvgRent = mean(RentPrice, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(Year = year(Year))

head(avg_rent)

p1 <- ggplot(avg_rent, aes(x = Year, y = AvgRent, color = County)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(
    title = "NYC Rental Prices within ZIP codes for all available years",
    x = "Year",
    y = "Rental Prices ($)",
    color = "County"
  ) +
  theme_minimal() +
    scale_x_continuous(
    breaks = seq(min(avg_rent$Year, na.rm = TRUE),
                 max(avg_rent$Year, na.rm = TRUE),
                 by = 1) 
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p1

```

The line chart shows the average rental prices across New York City boroughs over time. A few key patterns stand out: 1.Rental prices generally increased from 2015 to 2024 across all boroughs, reflecting rising housing demand and inflationary pressures in the NYC market. 2.Around 2020–2021, a temporary decline or plateau appears, likely corresponding to the COVID-19 pandemic’s impact on the rental market as urban migration slowed and vacancy rates increased.

### Average rental price within each ZIP code over each month in 2023

```{r}
rent2023 <- z_together %>%
  mutate(Year = year(as.Date(Date)))%>%
  filter(Year == 2023) %>%
  group_by(County, ZipCode) %>%
  summarize(
    AvgRent2023 = mean(RentPrice, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  drop_na(AvgRent2023)

library(tidyverse)
library(lubridate)

# 计算2023年各邮政编码的月平均租金
rent_2023 <- z_together %>%
  mutate(Year = year(as.Date(Date))) %>%
  filter(Year == 2023) %>%
  group_by(County, ZipCode) %>%
  summarize(
    AvgRent2023 = mean(RentPrice, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  drop_na(AvgRent2023)

# Plotting the distribution of rents across administrative districts
p2 <- ggplot(rent_2023, aes(x = County, y = AvgRent2023, fill = County)) +
  geom_violin(trim = FALSE, alpha = 0.7) +
  labs(
    title = "Distribution of ZIP-Code-Level Rental Prices Across NYC Boroughs (2023)",
    x = "Borough",
    y = "Average Monthly Rent ($)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1)
  )

p2
```

### Combine the two previous plots

```{r}
combined_plot <- p1 / p2
combined_plot
ggsave("results/combined_plot.png", combined_plot, width = 12, height = 10, dpi = 300)
```

## Problem 3

```{r}
# Demographic_data
demo_raw <- read_csv("demographic_data.csv") 
colnames(demo_raw) <- demo_raw[4, ]     
demo <- demo_raw[-c(1:4), ] %>% 
  janitor::remove_empty(which = "cols") %>%
  mutate(
    SEQN = as.integer(SEQN),
    sex = recode(sex,
                 "1" = "Male",
                 "2" = "Female"),
    age = as.numeric(age),
    BMI = as.numeric(BMI),
    education = recode(education,
                       "1" = "Less than high school",
                       "2" = "High school equivalent",
                       "3" = "More than high school")
  )  %>% 
  filter(age >= 21) %>% 
  drop_na() %>% 
  clean_names() %>% 
  select(-bmi)  

# Accelerometer_data
accel <- read_csv("accelerometer_data.csv") %>% 
  clean_names() %>% 
  pivot_longer(
    cols = starts_with("min"),  
    names_to = "minute",     
    values_to = "value"
  ) %>%
  mutate(
    minute = as.integer(str_remove(minute, "min"))
  )%>% 
  drop_na()

# merge together
acc_together<-accel %>%
  left_join(demo, by = c("seqn" = "seqn"))%>% 
  drop_na()
```

### Table: the number of men and women in each education category

```{r}
gender_edu_table <- demo %>%
  group_by(education, sex) %>%
  summarize(count = n(), .groups = "drop") %>%
  pivot_wider(names_from = sex, values_from = count)
gender_edu_table
```

Overall, the largest cohort comprises individuals with higher educational attainment (‘More than high school’), indicating a significant proportion of highly educated participants within the sample. Males are more prevalent in the secondary education group (high school equivalent), while females slightly outnumber males in the higher education group. This may reflect gender disparities across different educational levels within the sample.

### Total activities (y-axis) against age (x-axis) plot
```{r}
total_activity <- acc_together %>% 
  group_by(seqn, sex, age, education) %>%  
  summarize(total_activity = sum(value, na.rm = TRUE), .groups = "drop")
# Plot
ggplot(total_activity, aes(x = age, y = total_activity, color = sex)) +
  geom_point(alpha = 0.6) +           
  geom_smooth(method = "loess", se = FALSE) +  
  facet_wrap(~ education, scales = "free_y") + 
  labs(
    title = "Total Daily Activity vs Age by Sex and Education Level",
    x = "Age",
    y = "Total Activity",
    color = "Sex"
  ) +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))
```
Trend lines indicate that overall activity levels generally decline with age, though the rate of decline may differ between men and women.

### 
```{r}
hourly_activity <- acc_together %>%
  mutate(hour = floor(as.numeric(gsub("min", "", minute)) / 60)) %>% 
  group_by(sex, education, hour) %>%
  summarize(avg_activity = mean(value, na.rm = TRUE), .groups = "drop")

ggplot(hourly_activity, aes(x = hour, y = avg_activity, color = sex)) +
  geom_line(size = 1) +            
  geom_smooth(method = "loess", se = FALSE, linetype = "dashed") + 
  facet_wrap(~ education, ncol = 1) + 
  labs(
    title = "Average 24-hour Activity Curves by Education Level and Sex",
    x = "Hour of the Day",
    y = "Average Activity",
    color = "Sex"
  ) +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))
```
The trends observed in both males and females over a 24-hour period are similar. Levels decline between 0 and 5 hours, rise between 5 and 10 hours, fluctuate evenly around 10 to 18 hours, and then decline again between 18 and 24 hours.
